---
title: "score_f15_t4_spikeT3_band_0.4_minimap"
author: "youyupei"
date: "2020-01-27"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

```{r palette}
blue = "#6B93A7"
pink = "#CC5963"
green = "#289680"
orange = "#F4C300"
```


# Overall accuraccy 
## minimap2 mapping result
```{r CONSTANT}

library(ggplot2)
library(reshape2)
library(magrittr)
source("/home/ubuntu/PhD_proj/script_bin//helper.R")
source("~/PhD_proj/old_pipeline/Script/Probabilistic_model.R")




#get_likelihood_matrix = function(filename){
filename = "/home/ubuntu/PhD_proj/pipeline_1.21/Validation/score_f15_t4_spikeT3_band_0.4_new_likelihood_minimap.csv"



label_fast5_name <- function(f_names) {
  new_f_names = f_names
  for (name in unique(f_names)){
    num_rep = sum(f_names == name)
    new_name = paste0(name,'_',1:num_rep)
    new_f_names[which(f_names == name)] = new_name
        
  }
  return(new_f_names)
  
}

ncol = 50
  score_f15_t4 = read.csv(file = filename,header = F,col.names = seq(1, ncol))
  while(sum(is.numeric(score_f15_t4$X2)) > 0 || sum(is.na(score_f15_t4$X2))) {
    ncol = ncol + 50
    score_f15_t4 = read.csv(file = filename,header = F,col.names = seq(1, ncol))
  }
  fast5_filenames = as.character(score_f15_t4$X2)
  fast5_filenames = label_fast5_name(fast5_filenames)
  minimap_correctness = score_f15_t4$X1 
  # likelihood (single candidate junctions removed)
  likelihood = cbind(minimap_correctness, score_f15_t4[,seq(6,ncol,3)])
  fast5_filenames = fast5_filenames[!is.na(likelihood$X6)]
  likelihood = likelihood[!is.na(likelihood$X6),]
  
  has_multi_candidates = apply(likelihood[,-1],1,function(x) sum(!is.na(x)) > 1)
  likelihood = likelihood[has_multi_candidates,]
  fast5_filenames = fast5_filenames[has_multi_candidates]

# Threshold for removing similar candidate from null distribution (e.g. 0.8 means 0.8 of the alternative one)
NULL_PROB_THRES = 0.8


###################minimap accuracy#######################
print("minimap2 accuracy:")
mean(likelihood$minimap_correctness)
print("Number of tested squiggles:")
length(likelihood$minimap_correctness)
```


## NanoSplicer result

**Setting:**
Flank Size:15 bases in each side
Candidate Squiggle trim: 4 bases in each side
Tombo resquiggle: Correct squins isoform sequences provided.
Normalisation: global normalisation parameter from tombo.
Distance metrics in DTW: Negative log normal density
Outlier threshold: $\pm 3 \times MAD$

```{r Load File}
# distance of best matched candidate
min_dist = apply(likelihood[,-1],1,function(x) min(x, na.rm=T))
# bool that the corresponding prodiction is correct or not
correctness = apply(likelihood[,-1],1,which.min) == 1

print("NanoSplicer accuraccy per squiggle:")
print(sum(correctness)/length(correctness))
length(correctness)
```


# Probabilistic output
## Call z = k given H = 1
In this section, we used the spike-in dataset to simulate H=1, we compared the accuracy of picking the best as well as similar P(z)

```{r} 
thres = seq(0.0,1,0.01)

correctness0 = apply(likelihood[likelihood$minimap_correctness == 0,-1],1,function(x) {
  x = na.removed(x)
  true_value = x[1]
  min_value = min(x)
  abs(min_value-true_value) <= -log(thres)
  })

correctness1 = apply(likelihood[likelihood$minimap_correctness == 1,-1],1,function(x) {
  x = na.removed(x)
  true_value = x[1]
  min_value = min(x)
  abs(min_value-true_value) <= -log(thres)
  })



prop_scale = dim(correctness0)[2] + dim(correctness1)[2]
ggplot() + 
  theme(axis.text=element_text(size=12),plot.margin = margin(.0,2.8,.0,2.8, "cm")) +
  # overall acc
  geom_line(aes(x = thres, y = apply(cbind(correctness0,correctness1),1,sum)), color = pink, size = 1.5) +
  # total # of tested squiggles
  geom_hline(yintercept=dim(correctness0)[2] + dim(correctness1)[2], linetype="dashed", color = "black",size = 0.5) +
  # minimap2 correct
  geom_hline(yintercept=dim(correctness1)[2], linetype="dashed", color = blue,size = 0.9) +
  scale_y_continuous(sec.axis = sec_axis(~./(dim(correctness0)[2] + dim(correctness1)[2]), name = "")) + 
  xlab("Threshold ratio") +
  ylab("") +
  xlim(1,0.8) +
  ggtitle("") +
  scale_y_continuous(limits = c(8000,11600),sec.axis = sec_axis(~./(dim(correctness0)[2] + dim(correctness1)[2]), name = ""))

 ggplot() + 
  geom_line(aes(x = thres, y = apply(correctness1,1,sum)),size = 1.5,color = green) +
  geom_hline(yintercept=dim(correctness0)[2] + dim(correctness1)[2], linetype="dashed", color = "black",size = 0.5) +
  geom_hline(yintercept=dim(correctness1)[2], linetype="dashed", color = blue,size = 0.9) +
  scale_y_continuous(sec.axis = sec_axis(~./(dim(correctness0)[2] + dim(correctness1)[2]), name = "")) + 
  xlab("") +
  ylab("") +
  xlim(1,0.8) +
  ggtitle("") +
  theme(axis.text=element_text(size=12),plot.margin = margin(.0,2.8,.0,2.8, "cm")) +
  scale_y_continuous(limits = c(8000,9800),sec.axis = sec_axis(~./prop_scale, name = ""))

ggplot() + 
  geom_line(aes(x = thres, y = apply(correctness0,1,sum)),size = 1.5,color = green) +
  geom_hline(yintercept=dim(correctness0)[2], linetype="dashed", color = "black",size = 0.5) +
  geom_hline(yintercept=0, linetype="dashed", color = blue,size = 0.9) +
  scale_y_continuous(sec.axis = sec_axis(~./(dim(correctness0)[2] + dim(correctness1)[2]), name = "")) + 
  xlab("") +
  ylab("") +
  xlim(1,0.8) +
  ggtitle("") +
  theme(axis.text=element_text(size=12),plot.margin = margin(.0,2.8,.0,2.8, "cm")) +
    scale_y_continuous(limits = c(0,1800),sec.axis = sec_axis(~./dim(correctness0)[2], name = ""))
```



## H: whether or not the candidates container the truth

Measure a (un)certainty on whether candidates contain a splice site for Xi (i-th squiggle)

### Distribution of S|H
**S : maximum similarity between Xi and all candidates**
This are the empirical distributions for S|H=1 and S|H=0, which is from the density estimation:
  * distribution for S|H=1: density of S from the spike-in dataset analysis result, in which the ground truth is known.
  * distribution for S|H=0: density of S from the same result but the true candidates (known) have been ignored.
```{r, warning=FALSE}
alternative_vector = apply(likelihood[,c(-1)],1, min, na.rm = T)

best_false =  apply(likelihood[,c(-1,-2)],1,min,na.rm = T)

sec_best_candidate = apply(likelihood[,c(-1)],1,function(x) sort(x)[2])

filter_thres = NULL_PROB_THRES

best_false_filtered = apply(likelihood[,-1],1,function(x) {
  x = x[!is.na(x)]
  true_value = x[1]
  false_value = sort(x[-1])
  false_value[abs(false_value-true_value) > -log(filter_thres)][1]
  })

best_false_filtered = na.removed(best_false_filtered)
# remove candidates that very similar to the best one 
sec_best_filtered = apply(likelihood[,-1],1,function(x){
  x = x[!is.na(x)]
  x = sort(x)
  best = x[1]
  others = x[-1]
  others[abs(others - best) > -log(filter_thres)][1]
  })

sec_best_filtered = na.removed(sec_best_filtered)

#################################################################################
#########################          density  plot           ######################
#################################################################################
ggplot() +
  geom_density(aes(x = -alternative_vector,fill = "best candidates (Empirical H=1 distribution)"), alpha = 0.7)+
  geom_density(aes(x = -best_false_filtered, fill = "best false candidates \n (Estimated H=0 distribution from spike-ins)"), alpha = 0.6) +
  ggtitle("Different empirical distributions") +
  xlab("-loglikelihood")  + 
  ylab("Density") +
  xlim(-3,1)+
    theme(axis.text=element_text(size=15)) +
    scale_fill_manual(values = c(pink,blue))
```



### Test reliability of P(H=1|S) estimation

In this section, I randomly selected half of the squiggles, and removed the true candidate result for each of them to simulate the situation that $H_i = 0$. The full result set was splitted into a "True set" where  $H_i = 1$ and a "False set" where  $H_i = 0$. 

```{r}

# df_1: score distribution for true candidates
df_1 = approxfun(density(alternative_vector,na.rm = T))


df_0 <- approxfun(density(best_false))
p_post_best_false = apply(likelihood[,-1], 1,
                 function(x){
                   report_probability(v = x[!is.na(x)], df_0 = df_0,df_1 = df_1)$p_post
                 }
)

df_0 <- approxfun(density(best_false_filtered))
p_post_best_false_filtered = apply(likelihood[,-1], 1,
                 function(x){
                   report_probability(v = x[!is.na(x)], df_0 = df_0,df_1 = df_1)$p_post
                 }
)

df_0 <- approxfun(density(sec_best_candidate))
p_post_sec_best = apply(likelihood[,-1], 1,
                 function(x){
                   report_probability(v = x[!is.na(x)], df_0 = df_0,df_1 = df_1)$p_post
                 }
)


df_0 <- approxfun(density(sec_best_filtered))
p_post_sec_best_filtered = apply(likelihood[,-1], 1,
                 function(x){
                   report_probability(v = x[!is.na(x)], df_0 = df_0,df_1 = df_1)$p_post
                 }
)

df_0 <- approxfun(density(best_false_filtered))
p_post_emp_H0 = apply(likelihood[,c(-1,-2)], 1,
                 function(x){
                   report_probability(v = x[!is.na(x)], df_0 = df_0,df_1 = df_1)$p_post
                 }
)

df_0 <- approxfun(density(sec_best_filtered))
p_post_est_H0 = apply(likelihood[,c(-1,-2)], 1,
                 function(x){
                   report_probability(v = x[!is.na(x)], df_0 = df_0,df_1 = df_1)$p_post
                 }
)

rm(df_0)
#################################################################################
#########################    plot           ######################
#################################################################################


#simulation
false_index = sample(1:dim(likelihood)[1],floor(dim(likelihood)[1]/2))
count_0 = hist(p_post_emp_H0[false_index], breaks=12, plot=FALSE)
count_1 = hist(p_post_best_false_filtered[-false_index], breaks=count_0$breaks, include.lowest=TRUE, plot=FALSE)$counts

# distribution
ggplot() +
    theme(axis.text=element_text(size=15))+
    geom_density(aes(x =c(p_post_emp_H0[false_index],p_post_best_false_filtered[-false_index]),fill = "Best False"),alpha = 0.6,fill= green, color = green) + 
    xlab("")

# prop of H = 1
ggplot() +
  theme(axis.text=element_text(size=15))+
    geom_col(aes(y = count_1/(count_1+count_0$counts),x = count_0$mids),fill = pink,alpha = 0.5) +
    xlab("")+
    ylab("") + 
    ylim(0,1) + 
  xlim(0,1)
# prop of H =  0 
ggplot() +
  theme(axis.text=element_text(size=15))+
  geom_col(aes(y = 1 - count_1/(count_1+count_0$counts),x = count_0$mids),fill = blue,alpha = 0.5) +
  xlab("")+
  ylab("")+
  ylim(0,1) +
  xlim(0,1)


```

### Relationship between the accuracy and the P(H=1 | S)

```{r}
# df_1: score distribution for true candidates
df_1 = approxfun(density(alternative_vector,na.rm = T))
correctness = apply(likelihood[,-1],1,which.min) == 1

df_0 <- approxfun(density(best_false_filtered))


#################################################################################
#########################          density  plot           ######################
#################################################################################
p_post_plot = apply(likelihood[(likelihood$minimap_correctness == 1) * (correctness == 1) == 1,-1], 1,
                 function(x){
                   report_probability(v = x[!is.na(x)], df_0 = df_0,df_1 = df_1)$p_post
                 }
)


ggplot() +
    geom_density(aes(x = p_post_plot, fill = "1"),alpha = 0.5) + 
  # configure
  ggtitle("Both minimap2 and NanoSplicer correct") +
  xlab("")  + 
  ylab("") + theme(plot.title = element_text(hjust = 0.5),axis.text=element_text(size=12))  +
  scale_fill_manual(values = c(blue))

#################################################################################
#########################          density  plot           ######################
#################################################################################
    p_post_plot = apply(likelihood[(likelihood$minimap_correctness == 1) * (correctness == 0) == 1,-1], 1,
                 function(x){
                   report_probability(v = x[!is.na(x)], df_0 = df_0,df_1 = df_1)$p_post
                 }
)

ggplot() +
    geom_density(aes(x = p_post_plot, fill = "1"),alpha = 0.5) + 
  # configure
  ggtitle("Both minimap2 found but NanoSplicer missed") +
  xlab("")  + 
  ylab("") + theme(plot.title = element_text(hjust = 0.5),axis.text=element_text(size=12))  +
  scale_fill_manual(values = c(blue))

#################################################################################
#########################          density  plot           ######################
#################################################################################
p_post_plot = apply(likelihood[(likelihood$minimap_correctness == 0) * (correctness == 0) == 1,-1], 1,
                 function(x){
                   report_probability(v = x[!is.na(x)], df_0 = df_0,df_1 = df_1)$p_post
                 }
)

ggplot() +
    geom_density(aes(x = p_post_plot, fill = "1"),alpha = 0.5) + 
  # configure
  ggtitle("Both missed") +
  xlab("")  + 
  ylab("") + theme(plot.title = element_text(hjust = 0.5),axis.text=element_text(size=12))  +
  scale_fill_manual(values = c(blue))

#################################################################################
#########################          density  plot           ######################
#################################################################################
p_post_plot = apply(likelihood[correctness == 0,-1], 1,
                 function(x){
                   report_probability(v = x[!is.na(x)], df_0 = df_0,df_1 = df_1)$p_post
                 }
)

ggplot() +
    geom_density(aes(x = p_post_plot, fill = "1"),alpha = 0.5) + 
  # configure
  ggtitle("NanoSplcier wrong") +
  xlab("")  + 
  ylab("") + theme(plot.title = element_text(hjust = 0.5),axis.text=element_text(size=12))  +
  scale_fill_manual(values = c(blue))


#################################################################################
#########################          density  plot           ######################
#################################################################################
p_post_plot = apply(likelihood[,-1], 1,
                 function(x){
                   report_probability(v = x[!is.na(x)], df_0 = df_0,df_1 = df_1)$p_post
                 }
)

ggplot() +
    geom_density(aes(x = p_post_plot, fill = "1"),alpha = 0.5) + 
  # configure
  ggtitle("ALL") +
  xlab("")  + 
  ylab("") + theme(plot.title = element_text(hjust = 0.5),axis.text=element_text(size=12))  +
  scale_fill_manual(values = c(blue))

```

## P(H|S) and accuracy
```{r}
Q = p_post_plot
correctness
plot(y= correctness, x = Q, pch = 0.000001)

thres = seq(0,1,0.01)
y = sapply(thres, function(x){
  mean(correctness[Q<x])
})

ggplot() +
  geom_line(aes(y = y, x = thres), col = pink,size = 2) + 
  theme(plot.margin = margin(2,2.8,2,2.8, "cm"))

```

## Filter H


```{r}
p_cutoff = seq(0,0.99,0.01)
p_post = p_post_best_false_filtered
likelihood_0 = likelihood
correctness = sapply(p_cutoff,function(t){
    apply(likelihood_0[p_post > t ,-1],1,function(x){
      x = na.removed(x)
      which.min(x) == 1
    })
}
)

minimap_acc = sapply(p_cutoff,function(t){
   mean(likelihood[p_post > t ,]$minimap_correctness)
    })


accuracy = unlist(lapply(correctness,mean))
remain = unlist(sapply(p_cutoff,function(x) mean(p_post>x)))
ggplot() +
  theme(axis.text=element_text(size=15))+
  geom_line(aes(x = p_cutoff, y=accuracy,col = "NanoSplicer Accuracy"),size = 1) +
  geom_line(aes(x = p_cutoff, y=minimap_acc,col = "Minimap2 Accuracy"),size = 1) +
    geom_line(aes(x = p_cutoff, y=remain, col = "Prop of remaining squiggles after filtering"), linetype = "dashed",size = 1) +
  ylim(0.4,1) + xlim(0,1)+

  ylab("Accuraccy of choosing candidate \nwith largest P(z=k|x) as output") + 
  xlab("cutoff of P(H=1|x)")
```




